\papertitle

### 2014-03-20 final-report/paper.tex
Typos and new section

@@

The \fragment view comprises two charts: the utilization chart and the
operators chart. The utilization chart (Fig.~\ref{fig:utilization_chart}) shows on
how many workers a fragment is running over time. The chart is used to
quickly reveal any patterns in the schedule, like tail latency, long periods of
idleness, etc. We implemented a small brush (at the top of the figure) that
allows the user to zoom-in and further analyze the problematic area.

@@

The \overall shows small multiples of the utilization chart as described in the previous section.
Execution skew can be spotted in this view. Furthermore, the user can compare the fragments and find
correlations between the execution of different fragments. This view helps the user navigate to the
right fragment and is not intended to offer further inside into the causes or performance problems.

@@

@@

We explore a simple join query we performed on one million tuples extracted from the Twitter user graph. This query joins Twitter followers and followees by user ID. The query can be written in SQL as expressed below.

@@

%dom

@@

The \network visualization at a glance shows how much data was sent between workers by one query fragment to the next. The two \network views in Figure~\ref{fig:no-skew} and Figure~\ref{fig:skew} were obtained by selecting the Frag0$\rightarrow$Frag2 and Frag1$\rightarrow$Frag2 edges respectively in the \graph view. Figure~\ref{fig:network} displays aggregate data communication between workers. In Figure~\ref{fig:no-skew} the amount of data sent from each worker to every other worker is balanced as opposed to Figure~\ref{fig:skew} where one can see that the volume of data sent to worker~1 and worker~5 outweights data sent to the other workers in the system. This apparent data skew could be caused by partitioning issues.

@@

The \fragment visualization allows the user to dive down into query execution details at the worker level. The \fragment view in Figure~\ref{fig:exec_skew} was obtained by selecting Frag2, which is the fragment that performs the join, in the \graph view. The user can observe at a glance that certain workers produce much more data as a result of the join execution. Particularly, we saw in Case 2 Figure~\ref{fig:skew}, that tuples sent to Fragment 2 from Fragment 1 were heavily skewed towards being sent to worker~1 and worker~5. In Figure~\ref{fig:exec_skew}, the user can now see that as a result of the join, worker~5 ends up producing most tuples that have to be written back to disk. The color scheme used for each worker schedules in the \fragment view is color-coded to match the operators in Figure~\ref{fig:join}. Consequently, one can quickly observe in the \fragment view that worker~5 is busy sending join results from t=2.5s to t=5.0s, which differes from the rest of the workers. Using similar reasoning, one can conclude that worker 5 presents a performance bottleneck in the execution of the query on Myria. Perhaps the user can modify the way the input data set gets partitioned. Alternatively she could change the number of workers allocated for the job.

@@

All in all, \system allows the user to get much more insight into the execution of a query than what was available before, i.e. text-based performance logs, and runtimes. Being able to navigate performance data using a simple interface can indeed facilitate the task of a developper who is designing Myria's back end, or a Myria user who is writing complex queries to process large amounts of data. That said, \system can only help the user visualize query execution performance, but won't necessarily point the user to what she needs to do to improve performance of her query execution.

@@

This paper introduced \system, a visualization tool carefully tailored to help
database developers or users to quickly identify common distributed databases
performance bottlenecks, like tail latencies, data skew, etc. Our tool has been
successfully integrated with the Myria distributed database system, and has already
been used to identify several bugs, like poor physical query plan optimization,
tail latencies, poor data partitioning, which led us to believe the
visualization techniques we used are appropriate to help the user quickly
navigate to/find the cause of the performance problem. \system could be easily
plugged-in with another distributed database system, the only requirement being
a specific log data format.

>>>>

\todo{it's not a dag, as we noticed}

@@

The \fragment view comprises two charts: the utilization chart and the
operators chart. The utilization chart (Fig.~\ref{fig:utilization_chart}) shows on
how many workers a fragment is running over time. The chart is used to
quickly reveal any patterns in the schedule, like tail latency, long periods of
idleness, etc. Execution skew is visible as humps.
We implemented a small brush (at the top of the figure) that
allows the user to zoom-in and further analyze the problematic area.

@@

The \overall shows small multiples of the utilization chart as described in the previous section.
In addition to identifying execution skew, the user can compare the fragments and find
correlations between the execution of different fragments. This view helps the user navigate to the
right fragment and is not intended to offer further inside into the causes or performance problems.

@@

\todo{is this figure necessary or is the figure in the evaluation section enough?}

@@

We explore a simple join query we performed on one million tuples extracted from the Twitter user graph. This query joins Twitter followers and followees by user ID. The query can be written in SQL as:

@@

The \overall allows the user to quickly get an overview over the execution of all fragments
on all workers in the form of small multiples of the utilization charts as introduced in
Section~\ref{sec:fragment}. In Figure~\ref{fig:overview_skew} the user can see for example
that fragment~2 is very irregularly executed on different workers. If the graph is not a
single hill (as it is the case in Fragment 1, 3 and 4), workers are sleeping because they
either don't have enough data available to execute the operators or the output buffers
are full and data cannot be sent to the next fragment in the query graph.

The user has not identifies Fragment~2 to be a performance bottleneck and can use the other views
to further investigate its causes.

\begin{figure}[ht]
  \includegraphics[width=\columnwidth]{images/overview_skew}
  \caption{This \overall shows very irregular usage in fragment 2.}
  \label{fig:overview_skew}
\end{figure}

@@

\todo{Let's not call the fragments frag2 but write fragment 2}

The \network visualization at a glance shows how much data was sent between workers by one query fragment to the next. The two \network views in Figure~\ref{fig:network} were obtained by selecting the Frag0$\rightarrow$Frag2 and Frag1$\rightarrow$Frag2 edges respectively in the \graph view. Figure~\ref{fig:network} displays aggregate data communication between workers. In Figure~\ref{fig:no-skew} the amount of data sent from each worker to every other worker is balanced as opposed to Figure~\ref{fig:skew} where one can see that the volume of data sent to worker~1 and worker~5 outweighs data sent to the other workers in the system. This apparent data skew could be caused by partitioning issues.

@@

The \fragment visualization allows the user to dive down into query execution details at the worker level. The \fragment view in Figure~\ref{fig:exec_skew} was obtained by selecting Frag2, which is the fragment that performs the join, in the \graph view. The user can observe at a glance that certain workers produce much more data as a result of the join execution. Particularly, we saw in Case 2 Figure~\ref{fig:skew}, that tuples sent to Fragment 2 from Fragment 1 were heavily skewed towards being sent to worker~1 and worker~5. In Figure~\ref{fig:exec_skew}, the user can now see that as a result of the join, worker~5 ends up producing most tuples that have to be written back to disk. The color scheme used for each worker schedules in the \fragment view is color-coded to match the operators in Figure~\ref{fig:join}. Consequently, one can quickly observe in the \fragment view that worker~5 is busy sending join results from $t = 2.5s$ to $t = 5.0s$, which differs from the rest of the workers. Using similar reasoning, one can conclude that worker 5 presents a performance bottleneck in the execution of the query on Myria. In the short term, the user can modify the way the input data set gets partitioned. Alternatively she could change the number of workers allocated for the job. in the long term, developers should look into ways to improve data flow to avoid that workers have to wait for input data to process.

@@

All in all, \system allows the user to get much more insight into the execution of a query than what was available before, i.e. text-based performance logs, and runtimes. Being able to navigate performance data using a simple interface can indeed facilitate the task of a developer who is designing Myria's back end, or a Myria user who is writing complex queries to process large amounts of data. That said, \system can only help the user visualize query execution performance, but won't necessarily point the user to what she needs to do to improve performance of her query execution.

@@

This paper introduced \system, a visualization tool carefully tailored to help
database developers or users to quickly identify common distributed databases
performance bottlenecks, like tail latencies, data skew, etc. Our tool has been
successfully integrated with the Myria distributed database system, and has already
been used to identify several bugs, like poor physical query plan optimization,
tail latencies, poor data partitioning. We believe the visualization techniques
we used are appropriate to help the user quickly navigate to/find the cause of the
performance problem. \system could be easily plugged-in with another
distributed database system, the only requirement being a specific log data
format.
